---
title: "SLHW2"
author: "Ava Exelbirt, Sam Reade"
output: html_document
date: "2024-12-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
# Install necessary libraries if not installed
install.packages(c("ggplot2", "dplyr", "scales", "lubridate"))
install.packages("caret")
install.packages("GGally")
install.packages("tidyverse")
```


```{r libraries}
# Load libraries
library(ggplot2)
library(dplyr)
library(scales)
library(lubridate)
library(tidyverse)
library(GGally)
library(caret)
```

# About the Data

### Import Data


```{r data}
tuesdata <- tidytuesdayR::tt_load('2022-11-01')
tuesdata <- tidytuesdayR::tt_load(2022, week = 44)
horror <- tuesdata$horror_movies
```


```{r glimpse}
glimpse(horror)
```

### Data Dictionary

1.  The `id` variable is an integer that serves as a unique identifier for each movie.
2.  The `original_title` variable is a character string representing the movie's original title.
3.  The `title` variable is a character string containing the localized or alternative movie title.
4.  The `original_language` variable is a character field indicating the language in which the movie was originally made.
5.  The `overview` variable is a character field providing a brief description or synopsis of the movie.
6.  The `tagline` variable is a character field capturing the movie's catchphrase or slogan.
7.  The `release_date` variable is a date field that records the date when the movie was first released.
8.  The `poster_path` variable is a character field containing the URL to the movie's poster image.
9.  The `popularity` variable is a numerical value representing the movie's popularity score based on audience interactions.
10. The `vote_count` variable is an integer field that records the total number of audience votes received.
11. The `vote_average` variable is a numerical field that represents the average audience rating on a scale from 0 to 10.
12. The `budget` variable is an integer field capturing the movie's production budget in USD.
13. The `revenue` variable is an integer field indicating the total revenue earned by the movie in USD.
14. The `runtime` variable is an integer field that specifies the duration of the movie in minutes.
15. The `status` variable is a character field that indicates the current status of the movie, such as "Released."
16. The `adult` variable is a boolean that indicates whether the movie is intended for adult audiences.
17. The `backdrop_path` variable is a character field that provides the URL to the backdrop image for the movie.
18. The `genre_names` variable is a character field listing the genres associated with the movie, separated by commas.
19. The `collection` variable is a numerical field containing the unique ID of the collection the movie belongs to, which may be null for movies not part of a collection.
20. The `collection_name` variable is a character field representing the name of the collection, which may also be null if the movie does not belong to one.

### Available Data

The dataset contains detailed information on a wide range of horror movies, about \~35,000 pieces of entertainment, including various features such as title, genre, release date, runtime, popularity, budget, and revenue. Additional details include the movie's runtime, vote count, average vote, genre names, and collection association. Notably, the dataset also contains the poster and backdrop image URLs for each movie, as well as whether the movie is intended for adults. These data points provide a comprehensive view of each movie's performance, reception, and thematic elements, enabling further analysis on trends, movie popularity, and financial success within the horror genre. These features will be used to train a classification model to predict whether each entry is a successful movie or not. The objective is to leverage these data points to build an accurate classification model, focusing on identifying the key predictors that contribute most to the classification process.

### Motivation

As the entertainment industry expands, identifying the success of a movie is critical to content platforms and production companies. Success in the entertainment industry is typically measured by revenue, budget, and audience reception. Predicting whether a movie is likely to be successful or not can help guide investment decisions, optimize content strategies, and improve user recommendations. However, accurately predicting success is a challenge due to the multifaceted nature of what contributes to a movie's success, including budget, genre, release time, and audience engagement factors.

In this context, predicting a movie's success involves analyzing historical data and identifying patterns that correlate with positive outcomes. By doing so, production teams and platforms can better allocate resources, strategize marketing efforts, and predict the potential success of future movies. The motivation behind this project is to build a classification model that can predict whether a movie will be successful based on various features, thus improving decision-making processes in the entertainment industry.

### Goal

The primary goal of this project is to develop a classification model that predicts whether a given movie is successful or not. The project will focus on feature selection, model interpretation, and the comparison of predictor sets to determine the most significant factors contributing to a movie's success. By analyzing a variety of features such as budget, revenue, genre, and popularity, the goal is to build a model that classifies movies as "successful" or "unsuccessful" with high accuracy. This will allow content platforms and production companies to make data-driven decisions and better understand the elements that contribute to the success of a movie.

# Data Preprocessing and Visualization Tools

```{r summ}
summary(horror)
```

## Data Cleanup

### Handling NA Values

We will look at how many NA values are in each column to better understand our data set.

```{r NA-count}
na_counts <- colSums(is.na(horror))

print(na_counts)

na_counts_df <- data.frame(Column = names(na_counts), NA_Count = na_counts)
print(na_counts_df)

```
```{r}
sum(horror$revenue == 0)
sum(horror$budget == 0)
sum(horror$budget != 0 & horror$revenue != 0)
sum(horror$budget == 0 & horror$revenue == 0)

```



For numeric columns, we will fill missing values with the median values of that column. These include id, release_date, popularity, vote_count, vote_average, revenue, and runtime. We will then fill missing character columns with "Unknown." These include original_title, title, original_language, tagline, overview, poster_path, status, adult, and backdrop_path.

```{r NAs}
numeric_cols <- sapply(horror, is.numeric)
horror[numeric_cols] <- lapply(horror[numeric_cols], function(x) {
  ifelse(is.na(x), median(x, na.rm = TRUE), x)
})

## Fill missing character columns with "Unknown"
char_cols <- sapply(horror, is.character)
horror[char_cols] <- lapply(horror[char_cols], function(x) {
  ifelse(is.na(x), "Unknown", x)
})

```

### Drop Columns

We will remove the columns ids and paths as these are not needed for our overall analysis.

```{r drop}
horror <- horror %>%
  select(-c(id, poster_path, backdrop_path, collection, collection_name))
```

### Feature Engineering

As part of feature engineering we need to create our boolean-like columns to logical data types. We will do so for the adult column. If the observation is FALSE, then it will convert to a logical operator of 0. If the observation is TRUE for this column, then it will be converted to 1. We must also convert categorical columns to factors. This includes original_language, status, and genre_names. Finally, we will extract year from release_date because this will help in further analysis.

```{r feature-eng}
horror$adult <- as.logical(horror$adult)

categorical_cols <- c("original_language", "status", "genre_names")
horror[categorical_cols] <- lapply(horror[categorical_cols], as.factor)

horror$release_year <- as.numeric(substr(horror$release_date, 1, 4))
```

### Handling Outliers

We will replace some outliers. Specifically, for runtime we will replace runtime with the 0 if there is a runtime that is defined as an outlier, we will replace it with 0. We will also remove rows with outliers regarding popularity that is defined as popularity above 10000. We will also categorize budget levels. We categorize movies into "Low", "Medium", or "High" budget based on the budget column:

```{r outliers}
runQ1 <- quantile(horror$runtime, 0.25, na.rm = TRUE)
runQ3 <- quantile(horror$runtime, 0.75, na.rm = TRUE)
IQR <- runQ3 - runQ1
lower_bound <- runQ1 - 1.5 * IQR
upper_bound <- runQ3 + 1.5 * IQR
horror$runtime[horror$runtime < lower_bound | horror$runtime > upper_bound] <- 0

#horror$runtime[horror$runtime <= 0 | horror$runtime > 300] <- NA

horror <- horror[!(horror$popularity > 10000), ]

horror$budget_category <- ifelse(horror$budget == 0, "No Budget",
                           ifelse(horror$budget < 1e7, "Low",
                           ifelse(horror$budget < 5e7, "Medium", "High")))
```


### Create Target Variables

We first create a profit variable that is the revenue minus budget of a movie. We then create a success variable: if profit \> 0, movie is considered successful

```{r make-targ}
horror$profit <- horror$revenue - horror$budget
horror$success <- ifelse(horror$profit > 0, "Success", "No Success")

```

### Correlation Analysis

```{r corr}
ggcorr(horror[ , sapply(horror, is.numeric)], label = TRUE)
```

## Visualization Tools

### Distribution of Numeric Features

We will plot the distributions of numeric features, specifically budget, revenue, and runtime.

```{r num-dist}
ggplot(horror, aes(x = budget)) +
  geom_histogram(binwidth = 1e7, fill = "blue", color = "black", alpha = 0.7) +
  labs(title = "Distribution of Budget", x = "Budget", y = "Count") +
  theme_minimal()

ggplot(horror, aes(x = revenue)) +
  geom_histogram(binwidth = 1e7, fill = "green", color = "black", alpha = 0.7) +
  labs(title = "Distribution of Revenue", x = "Revenue", y = "Count") +
  theme_minimal()

ggplot(horror, aes(x = runtime)) +
  geom_histogram(binwidth = 10, fill = "red", color = "black", alpha = 0.7) +
  labs(title = "Distribution of Runtime", x = "Runtime (Minutes)", y = "Count") +
  theme_minimal()
```

```{r}
# For Budget (which has large values)
ggplot(horror, aes(x = budget)) +
  geom_histogram(binwidth = 1e8, fill = "blue", color = "black", alpha = 0.7) +  # Increase binwidth
  labs(title = "Distribution of Budget", x = "Budget", y = "Count") +
  theme_minimal()

# For Revenue (which also has large values)
ggplot(horror, aes(x = revenue)) +
  geom_histogram(binwidth = 1e8, fill = "green", color = "black", alpha = 0.7) +  # Increase binwidth
  labs(title = "Distribution of Revenue", x = "Revenue", y = "Count") +
  theme_minimal()

# For Runtime (which typically has smaller values)
ggplot(horror, aes(x = runtime)) +
  geom_histogram(binwidth = 5, fill = "red", color = "black", alpha = 0.7) +  # Adjust binwidth for runtime
  labs(title = "Distribution of Runtime", x = "Runtime (Minutes)", y = "Count") +
  theme_minimal()

```


Budget and revenue are both right skewed and unimodal. Runtime is also right skewed, but bimodal.

### Distribution of Some Categorical Features

We will plot the distributions of some categorical features, specifically release_year and budget_category.

```{r cat-dist}
ggplot(horror, aes(x = release_year)) +
  geom_bar(fill = "purple", color = "black") +
  labs(title = "Distribution of Movies by Release Year", x = "Release Year", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggplot(horror, aes(x = budget_category, fill = budget_category)) +
  geom_bar() +
  labs(title = "Distribution of Movies by Budget Category", x = "Budget Category", y = "Count") +
  theme_minimal() +
  theme(legend.position = "none")

```

Release year is left skewed and unimodal.

### Distribution of Target Variable

We will now look at the distribution of our target variable, show_or_movie.

###CHANGE WITH NEW PREDICTOR

```{r}
sum(horror$success == "Success")
sum(horror$success == "No Success")
```


```{r horrordist}
ggplot(horror, aes(x = success)) + 
  geom_bar(fill = "lightblue", color = "black") + 
  labs(title = "Distribution of Success of a Movie", x = "Success", y = "Count") + 
  theme_minimal()
```

### Variable Relationships

We first examine the relationship between budget and revenue for each horror movie, with points colored by their budget category, helping to identify patterns and outliers in how budget impacts revenue. We can also visualize how the budget has evolved over the years by plotting release_year versus budget. We then examine the relationship between `popularity` and `vote_average` to see if there's a trend in how movies' popularity correlates with their ratings. We also examine the relationship between `budget` and `profit`. Each point represents a movie, and the color differentiates between successful and non-successful movies. The idea is to see if movies with higher budgets tend to generate more profit. Finally we show how the `profit` distribution differs between successful and non-successful movies. This boxplot shows the distribution of `profit` for movies categorized as "Success" or "No Success". The plot helps visualize how profits are distributed across these categories, revealing if successful movies tend to have higher profits.

```{r}
ggplot(horror, aes(x = budget, y = revenue)) +
  geom_point(aes(color = budget_category), alpha = 0.7) +
  scale_x_continuous(labels = scales::comma) +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "Budget vs Revenue", x = "Budget", y = "Revenue") +
  theme_minimal() +
  theme(legend.title = element_blank())

ggplot(horror, aes(x = release_year, y = budget)) +
  geom_point(aes(color = budget_category), alpha = 0.7) +
  scale_x_continuous(labels = scales::comma) +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "Release Year vs")

ggplot(horror, aes(x = popularity, y = vote_average)) +
  geom_point(aes(color = genre_names), alpha = 0.7) +
  scale_x_continuous(labels = scales::comma) +
  labs(title = "Popularity vs Vote Average", x = "Popularity", y = "Vote Average") +
  theme_minimal() +
  theme(legend.position = "none")

ggplot(horror, aes(x = budget, y = profit)) + 
  geom_point(aes(color = success), alpha = 0.6) + 
  scale_x_continuous(labels = scales::comma) + 
  scale_y_continuous(labels = scales::comma) + 
  labs(title = "Budget vs. Profit", x = "Budget", y = "Profit") + 
  theme_minimal() + 
  theme(legend.title = element_blank())

ggplot(horror, aes(x = success, y = profit, fill = success)) + 
  geom_boxplot() + 
  scale_y_continuous(labels = scales::comma) + 
  labs(title = "Profit Distribution by Success", x = "Success", y = "Profit") + 
  theme_minimal() + 
  theme(legend.title = element_blank())

```

```{r}
sum(horror$budget == 0)
sum(horror$budget != 0)
```



## Split the Data

### Train and Test Data

We will split the data into training (60%) and testing (40%) sets. We will then look at the new data by checking the number of rows in training and testing sets and looking at the summary of the training set.

```{r split}
set.seed(123)

in_train <- createDataPartition(horror$budget_category, p = 0.6, list = FALSE)
training <- horror[in_train, ]
testing <- horror[-in_train, ]

nrow(training)
nrow(testing)

summary(training)
```

### Distribution of Target Variable

```{r show-dist}
table(training$success) / length(training$success)
```

4% successful movies, \ ~96% unsuccessful movies: unbalanced dataset 

### Training Visuals

```{r}
# Visualizing the distribution of 'budget' in training data
ggplot(training, aes(x = budget)) + 
  geom_density(fill = "lightblue") + 
  labs(title = "Density of Budget", x = "Budget", y = "Density")

# Visualizing the distribution of 'revenue' in training data
ggplot(training, aes(x = revenue)) + 
  geom_density(fill = "lightgreen") + 
  labs(title = "Density of Revenue", x = "Revenue", y = "Density")

# Visualizing 'budget' against 'revenue'
ggplot(training, aes(x = budget, y = revenue)) + 
  geom_point(aes(color = budget_category), alpha = 0.7) + 
  scale_x_continuous(labels = scales::comma) + 
  scale_y_continuous(labels = scales::comma) + 
  labs(title = "Budget vs Revenue", x = "Budget", y = "Revenue") + 
  theme_minimal() + 
  theme(legend.title = element_blank())

# Exploring 'runtime' distribution
ggplot(training, aes(x = runtime)) + 
  geom_density(fill = "lightcoral") + 
  labs(title = "Density of Runtime", x = "Runtime", y = "Density")

# Exploring 'popularity' distribution
ggplot(training, aes(x = popularity)) + 
  geom_density(fill = "lightyellow") + 
  labs(title = "Density of Popularity", x = "Popularity", y = "Density")

# Decomposing runtime by 'budget_category'
ggplot(training, aes(x = runtime, fill = budget_category)) + 
  geom_density(alpha = 0.5) + 
  labs(title = "Runtime by Budget Category", x = "Runtime", y = "Density")

# Visualizing 'vote_average' distribution
ggplot(training, aes(x = vote_average)) + 
  geom_density(fill = "lightblue") + 
  labs(title = "Density of Vote Average", x = "Vote Average", y = "Density")

# Visualizing 'release_year' distribution
ggplot(training, aes(x = release_year)) + 
  geom_bar(fill = "lightgreen") + 
  labs(title = "Release Year Distribution", x = "Release Year", y = "Count")

# Visualizing the distribution of 'success' in training data
ggplot(training, aes(x = success)) + 
  geom_bar(fill = "lightblue", color = "black") + 
  labs(title = "Distribution of Success", x = "Success", y = "Count") + 
  theme_minimal()

```

# Classification with Emphasis on Prediction

# Questions: 
# "How well can we predict a movie's revenue based on its budget, popularity, 
#  and release year?" (Regression)

# Predicting popularity (Regression)

# "Can we predict whether a movie will be profitable?" (QDA and LDA)

# "predicting succes categories (hit, average, flop) (QDA and LDA)



# Predicting Budget using logistic regression

```{r}
# Create a new data frame without rows where revenue is 0
df_for_budg = subset(horror, revenue != 0)

sum(horror$revenue == 0)
```

```{r}
# Separate predictors (features) and target variable (budget)
features_for_budg <- horror[, c("revenue", "popularity", "release_year")]
target_budg <- horror$budget
```

# Visualizaing the Relationships

```{r}
library(ggplot2)

# Scatter plots to see the relationships
ggplot(df_for_budg, aes(x = revenue, y = budget)) +
  geom_point() + 
  labs(title = "Budget vs Revenue", x = "Revenue", y = "Budget") +
  theme_minimal()

ggplot(df_for_budg, aes(x = release_year, y = budget)) +
  geom_point() + 
  labs(title = "Budget vs Release Year", x = "Release Year", y = "Budget") +
  theme_minimal()

ggplot(df_for_budg, aes(x = popularity, y = budget)) +
  geom_point() + 
  labs(title = "Budget vs Popularity", x = "Popularity", y = "Budget") +
  theme_minimal()

```
```{r}
max(df_for_budg$budget)
max(df_for_budg$revenue)
mean(df_for_budg$budget)
mean(df_for_budg$revenue)
```



# Creating categories for logistic regression

```{r}
budget_bins <- cut(df_for_budg$budget, 
                   breaks = c(0, 10000000, 20000000, 50000000 , Inf), 
                   labels = c("Low", "Medium", "High", "Very High"))

df_for_budg$budget_category <- budget_bins
```

# Fitting the model

```{r}
library(nnet)

multi_logistic_model <- multinom(budget_category ~ revenue + release_year + popularity, 
                                 data = df_for_budg)

summary(multi_logistic_model)
```

# Predicting the categories

```{r}
# Predict the budget category for the new data
predicted_categories <- predict(multi_logistic_model, df_for_budg)

# Print predicted categories
print(predicted_categories)
```


# Evaluating the Model 

```{r}
predicted_train_categories <- predict(multi_logistic_model, df_for_budg)

table(Predicted = predicted_train_categories, Actual = df_for_budg$budget_category)

```


